---
title: "Project 1 Checkpoint 3"
format: html
editor: visual
---

```{r}
library(tidyverse)
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-26/plastics.csv')
```

## Data Summaries Using Existing Columns

### Data Summary 1

**Question**: Which parent companies contribute the most total plastic waste across all countries, and what share of the total recorded waste does each company account for?

```{r}
overall_total <- sum(plastics$grand_total, na.rm = TRUE)

company_share <- function(x, overall_total) {
  sum(x, na.rm = TRUE) / overall_total
}

company_summary <- plastics |>
  filter(!is.na(parent_company),
         parent_company != "Grand Total",
         parent_company != "null",
         parent_company != "NULL",
         parent_company != "Unbranded") |>
  group_by(parent_company) |>
  summarize(
    total_plastic_waste = sum(grand_total, na.rm = TRUE),
    share_of_total = company_share(grand_total, overall_total),
    .groups = "drop"
  ) |>
  arrange(desc(total_plastic_waste)) |>
  slice_head(n = 10)

company_summary
```

### Data Summary 2

**Question**: Which countries had clean up efforts that found the largest grand total of plastics in 2019 and 2020?

```{r}
## Uses iteration and a custom function
total_plastics <- function(input_country){
  
  grand_total_plastics <- plastics |>
    filter(country == input_country) |>
    distinct(year, grand_total) |>
    pull(grand_total)
  
  sum(grand_total_plastics, na.rm = TRUE)
}

plastics |>
  distinct(country) |>
  filter(country != "EMPTY") |>
  mutate(
    plastic_total = map_int(country, total_plastics)
    ) |>
  slice_max(plastic_total, n = 10)
```

## Data Summaries That Modify Columns

### Data Summary 3

**Question**: Which countries have the highest average amount of plastic collected per cleanup event?

```{r}
plastic_per_event <- function(input_country) {
  plastics |>
    filter(country == input_country) |>
    distinct(year, grand_total, num_events) |>
    summarise(
      total_plastic = sum(grand_total, na.rm = TRUE),
      total_events = sum(num_events, na.rm = TRUE)
    ) |>
    mutate(plastic_per_event = total_plastic / total_events) |>
    pull(plastic_per_event)
}

plastic_event_summary <- plastics |>
  distinct(country) |>
  mutate(plastic_per_event = map_dbl(country, plastic_per_event)) |>
  slice_max(plastic_per_event, n = 10)

plastic_event_summary
```

### Data Summary 4

**Question**: Which nations had the highest average participation for cleanup events in 2019 and 2020?

```{r}
cleanup_totals <- function(input_country) {
  plastics |>
    filter(country == input_country) |>
    distinct(num_events, volunteers) |>
    mutate(total_vol = sum(volunteers, na.rm = T),
         total_events = sum(num_events, na.rm = T)) |>
    distinct(total_vol, total_events) |>
    mutate(vol_per_event = total_vol / total_events) |>
    pull(vol_per_event)
}

plastics |>
  distinct(country) |>
  mutate(
    avg_participation = map_dbl(country, cleanup_totals),
    rounded_participation = round(avg_participation)
  ) |>
  slice_max(avg_participation, n = 20)
```

```{r}
## Renaming the countries with top 10 plastic totals
plastics |>
  distinct(country) |>
  filter(country != "EMPTY") |>
  mutate(
    plastic_total = map_int(country, total_plastics)
    ) |>
  filter(country != "Nigeria" & plastic_total != 61123) |>
  slice_max(plastic_total, n = 10) |> 
  mutate(country = recode(country, 
                          `Taiwan_ Republic of China (ROC)` = "Taiwan",
                          NIGERIA = "Nigeria", 
                          ECUADOR = "Ecuador",
                          `United States of America` = "USA",
                          `Burkina Faso` = "BFA"))
  
```

```{r}
library(glue)
library(httr)
library(tidyjson)
library(jsonlite)
library(readr)
library(tidyverse)
```

```{r}
my_key <- "eAqHuhzDSfLFeFP2pqhsbaC5iBQ6SvZ2VJeYSeNB"
## Create a function to pull the GDP of countries from the API
get_gdp <- function(country_code, key = my_key) {
  url <- glue("https://api.api-ninjas.com/v1/gdp?country={country_code};X-Api-Key={key}")
  
  my_data <- safely(fromJSON)(url)
  
  if (is.null(my_data$result)) {
    return(NA)
  } else {
    return(my_data$result |> filter(year %in% c(2019, 2020)))
  }
}
```

```{r}
## Create a list of the top 10 countries with the highest total plastics
library(countrycode)

top_plastic_countries <- plastics |>
  distinct(country) |>
  filter(!country %in% c("EMPTY", "Taiwan_ Republic of China (ROC)")) |>  ## Dropping Taiwan because of its contested status in the world (some supplemental data does not include Taiwan)
  mutate(
    plastic_total = map_int(country, total_plastics),
    country_code = countrycode(country, origin = "country.name", destination = "iso3c")
  ) |>
  filter(!is.na(country_code)) |> ## Dropping Nigeria's duplicate
  filter(plastic_total != 61123) |> ## Added security for removing the specific duplicate
  slice_max(plastic_total, n = 10) |>
  pull(country_code)
```

```{r}
#| eval: false
Sys.sleep(20)
## Create a list of the countries with their GDP
gdp_list <- top_plastic_countries |>
  map(get_gdp)

## Double check for NAs
gdp_list |>
  is.na() |>
  sum()
```

```{r}
#| eval: false
## Turn this into a tibble to join with our original data frame
gdp_data <- gdp_list |>
  keep(is.data.frame) |>
  bind_rows() |>
  rename(country_code = country)

gdp_data |>
  pull(country_code)

## Saving the csv locally
write.csv(gdp_data, "gdp.csv", row.names = FALSE)
```

```{r}
## Rereading in the data from the local file
gdp_df <- read.csv("gdp.csv")
```

```{r}
#| warning: false
plastics_top <- plastics |>
  filter(country != "Nigeria") |>
  filter(!country %in% c("EMPTY", "Taiwan_ Republic of China (ROC)")) |>
  mutate(
    country_code = countrycode(country, origin = "country.name", destination = "iso3c")
  ) |>
  filter(country_code %in% top_plastic_countries) |>
  left_join(gdp_df)

head(plastics_top)
```

```{r}
renewable_data <- read.csv("renewable-energy.csv")
## Data found from https://github.com/datasets/world-development-indicators/tree/main/indicators/eg.fec.rnew.zs

plastics_top <- plastics_top |>
  left_join(renewable_data |>
              filter(Year %in% c(2019, 2020),
                     Country.Code %in% plastics_top$country_code), 
            by = c("country_code" = "Country.Code", 
                   "year" = "Year")) |>
  select(-Country.Name)
```
