---
title: "Project 1 Checkpoint 3"
format: html
editor: visual
---

```{r}
library(tidyverse)
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-26/plastics.csv')
```

## Data Summaries Using Existing Columns

### Data Summary 1

**Question**: Which parent companies contribute the most total plastic waste across all countries, and what share of the total recorded waste does each company account for?

```{r}
overall_total <- sum(plastics$grand_total, na.rm = TRUE)

company_share <- function(x, overall_total) {
  sum(x, na.rm = TRUE) / overall_total
}

company_summary <- plastics |>
  filter(!is.na(parent_company),
         parent_company != "Grand Total",
         parent_company != "null",
         parent_company != "NULL",
         parent_company != "Unbranded") |>
  group_by(parent_company) |>
  summarize(
    total_plastic_waste = sum(grand_total, na.rm = TRUE),
    share_of_total = company_share(grand_total, overall_total),
    .groups = "drop"
  ) |>
  arrange(desc(total_plastic_waste)) |>
  slice_head(n = 10)

company_summary
```

### Data Summary 2

**Question**: Which countries had clean up efforts that found the largest grand total of plastics in 2019 and 2020?

```{r}
## Uses iteration and a custom function
total_plastics <- function(input_country){
  
  grand_total_plastics <- plastics |>
    filter(country == input_country) |>
    distinct(year, grand_total) |>
    pull(grand_total)
  
  sum(grand_total_plastics, na.rm = TRUE)
}

plastics |>
  distinct(country) |>
  filter(country != "EMPTY") |>
  mutate(
    plastic_total = map_int(country, total_plastics)
    ) |>
  slice_max(plastic_total, n = 10)
```

## Data Summaries That Modify Columns

### Data Summary 3

**Question**: Which countries have the highest average amount of plastic collected per cleanup event?

```{r}
plastic_per_event <- function(input_country) {
  plastics |>
    filter(country == input_country) |>
    distinct(year, grand_total, num_events) |>
    summarise(
      total_plastic = sum(grand_total, na.rm = TRUE),
      total_events = sum(num_events, na.rm = TRUE)
    ) |>
    mutate(plastic_per_event = total_plastic / total_events) |>
    pull(plastic_per_event)
}

plastic_event_summary <- plastics |>
  distinct(country) |>
  mutate(plastic_per_event = map_dbl(country, plastic_per_event)) |>
  slice_max(plastic_per_event, n = 10)

plastic_event_summary
```

### Data Summary 4

**Question**: Which nations had the highest average participation for cleanup events in 2019 and 2020?

```{r}
cleanup_totals <- function(input_country) {
  plastics |>
    filter(country == input_country) |>
    distinct(num_events, volunteers) |>
    mutate(total_vol = sum(volunteers, na.rm = T),
         total_events = sum(num_events, na.rm = T)) |>
    distinct(total_vol, total_events) |>
    mutate(vol_per_event = total_vol / total_events) |>
    pull(vol_per_event)
}

plastics |>
  distinct(country) |>
  mutate(
    avg_participation = map_dbl(country, cleanup_totals),
    rounded_participation = round(avg_participation)
  ) |>
  slice_max(avg_participation, n = 20)
```

```{r}
## Renaming the countries with top 10 plastic totals
plastics |>
  distinct(country) |>
  filter(country != "EMPTY") |>
  mutate(
    plastic_total = map_int(country, total_plastics)
    ) |>
  filter(country != "Nigeria" & plastic_total != 61123) |>
  slice_max(plastic_total, n = 10) |> 
  mutate(country = recode(country, 
                          `Taiwan_ Republic of China (ROC)` = "Taiwan",
                          NIGERIA = "Nigeria", 
                          ECUADOR = "Ecuador",
                          `United States of America` = "USA",
                          `Burkina Faso` = "BFA"))
  
```

```{r}
library(glue)
library(httr)
library(tidyjson)
library(jsonlite)
library(readr)
library(tidyverse)
```

```{r}
my_key <- "key"
## Create a function to pull the GDP of countries from the API
get_gdp <- function(country_code, key = my_key) {
  url <- glue("https://api.api-ninjas.com/v1/gdp?country={country_code};X-Api-Key={key}")
  
  my_data <- safely(fromJSON)(url)
  
  if (is.null(my_data$result)) {
    return(NA)
  } else {
    return(my_data$result |> filter(year %in% c(2019, 2020)))
  }
}
```

```{r}
## Create a list of the top 10 countries with the highest total plastics
library(countrycode)

top_plastic_countries <- plastics |>
  distinct(country) |>
  filter(!country %in% c("EMPTY", "Taiwan_ Republic of China (ROC)")) |>  ## Dropping Taiwan because of its contested status in the world (some supplemental data does not include Taiwan)
  mutate(
    plastic_total = map_int(country, total_plastics),
    country_code = countrycode(country, origin = "country.name", destination = "iso3c")
  ) |>
  filter(!is.na(country_code)) |> ## Dropping Nigeria's duplicate
  filter(plastic_total != 61123) |> ## Added security for removing the specific duplicate
  slice_max(plastic_total, n = 10) |>
  pull(country_code)
```

```{r}
#| eval: false
Sys.sleep(20)
## Create a list of the countries with their GDP
gdp_list <- top_plastic_countries |>
  map(get_gdp)

## Double check for NAs
gdp_list |>
  is.na() |>
  sum()
```

```{r}
#| eval: false
## Turn this into a tibble to join with our original data frame
gdp_data <- gdp_list |>
  keep(is.data.frame) |>
  bind_rows() |>
  rename(country_code = country)

gdp_data |>
  pull(country_code)

## Saving the csv locally
write.csv(gdp_data, "gdp.csv", row.names = FALSE)
```

**GDP Dataset Description**

The GDP data set has been extracted from the API Ninjas "GDP" API. In this case, the data has been programatically obtained in R by making an API call for each country by its ISO3 country code. Annual Gross Domestic Product (GDP) in U.S. Dollars has been recorded in this data set.

However, for this assignment, the years included have been limited to 2019 and 2020. Therefore, each country will appear twice in this data set corresponding to the two years. The data is in the form of a Tidy Data Frame, and it would later be merged with the plastic data set.

Key variables include:

-   country: ISO3 country code

<!-- -->

-   year: year of observation (2019 or 2020)

-   value: GDP (nominal) in U.S. dollars

```{r}
## Rereading in the data from the local file
gdp_df <- read.csv("gdp.csv")
```

```{r}
#| warning: false
plastics_top <- plastics |>
  filter(country != "Nigeria") |>
  filter(!country %in% c("EMPTY", "Taiwan_ Republic of China (ROC)")) |>
  mutate(
    country_code = countrycode(country, origin = "country.name", destination = "iso3c")
  ) |>
  filter(country_code %in% top_plastic_countries) |>
  left_join(gdp_df)

head(plastics_top)
```

```{r}
renewable_data <- read.csv("renewable-energy.csv")
## Data found from https://github.com/datasets/world-development-indicators/tree/main/indicators/eg.fec.rnew.zs

plastics_top <- plastics_top |>
  left_join(renewable_data |>
              filter(Year %in% c(2019, 2020),
                     Country.Code %in% plastics_top$country_code), 
            by = c("country_code" = "Country.Code", 
                   "year" = "Year")) |>
  select(-Country.Name)
```

**Renewable Energy Dataset Description**

The renewable energy dataset was sourced from an open-source data set provided on GitHub and is based on World Development Indicators. The data set consists of country-specific information about renewable energy consumption over time.

In this case, the renewable energy data set has been filtered to include only the years 2019 and 2020 as well as only those countries that appear in the plastic data set. This filtering was done using ISO3 codes.

Key variables include:

-   Country Code : ISO3 country code

-   Year: year of observation

-   renewable energy consumption, measured as the percentage of total final energy consumption from renewable sources

This dataset was merged with the GDP and plastics datasets to support comparative analysis across countries.

**GDP Data Summary**

```{r}
top10_codes <- plastics_top |>
  distinct(country_code) |>
  slice_head(n = 10) |>
  pull(country_code)

country_plot <- plastics_top |>
  filter(country_code %in% top10_codes) |>
  group_by(country, country_code) |>
  summarize(
    total_plastic = sum(grand_total, na.rm = TRUE),
    gdp = mean(gdp_nominal, na.rm = TRUE),
    .groups = "drop"
  )

ggplot(country_plot, aes(x = gdp, y = total_plastic)) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_text(aes(label = country_code), vjust = -0.5, size = 3) +
  labs(
    title = "Relationship Between GDP and Total Plastics (Top 10 Countries)",
    x = "GDP (billions USD)",
    y = "Total Plastics"
  ) +
  theme_minimal()
```
