---
title: "Project 1 Checkpoint 2 (Final Version)"
author: "Amanda Iang, Siddharth Venkiteswaran, Spencer Chinn"
format: html
editor: visual
---

## Project 1 Checkpoint 2

```{r}
library(tidyverse)
plastics <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-26/plastics.csv")
```

## Data Summaries Using Existing Columns

### Data Summary 1

**Question**: Which parent companies contribute the most total plastic waste across all countries, and what share of the total recorded waste does each company account for?

```{r}
overall_total <- sum(plastics$grand_total, na.rm = TRUE)

company_share <- function(x, overall_total) {
  sum(x, na.rm = TRUE) / overall_total
}

company_summary <- plastics |>
  filter(!is.na(parent_company),
         parent_company != "Grand Total",
         parent_company != "null",
         parent_company != "NULL",
         parent_company != "Unbranded") |>
  group_by(parent_company) |>
  summarize(
    total_plastic_waste = sum(grand_total, na.rm = TRUE),
    share_of_total = company_share(grand_total, overall_total),
    .groups = "drop"
  ) |>
  arrange(desc(total_plastic_waste)) |>
  slice_head(n = 10)

company_summary
```

### Data Summary 2

**Question**: Which countries had clean up efforts that found the largest grand total of plastics in 2019 and 2020?

```{r}
total_plastics <- function(input_country) {
  grand_total_plastics <- plastics |>
    filter(country == input_country) |>
    distinct(year, grand_total) |>
    pull(grand_total)

  sum(grand_total_plastics, na.rm = TRUE)
}

country_totals <- plastics |>
  distinct(country) |>
  mutate(plastic_total = map_int(country, total_plastics)) |>
  slice_max(plastic_total, n = 10)

country_totals
```

## Data Summaries That Modify Columns

### Data Summary 3

**Question**: Which countries have the highest average amount of plastic collected per cleanup event?

```{r}
plastic_per_event <- function(input_country) {
  plastics |>
    filter(country == input_country) |>
    distinct(year, grand_total, num_events) |>
    summarise(
      total_plastic = sum(grand_total, na.rm = TRUE),
      total_events = sum(num_events, na.rm = TRUE)
    ) |>
    mutate(plastic_per_event = total_plastic / total_events) |>
    pull(plastic_per_event)
}

plastic_event_summary <- plastics |>
  distinct(country) |>
  mutate(plastic_per_event = map_dbl(country, plastic_per_event)) |>
  slice_max(plastic_per_event, n = 10)

plastic_event_summary
```

### Data Summary 4

**Question**: Which nations had the highest average participation for cleanup events in 2019 and 2020?

```{r}
cleanup_totals <- function(input_country) {
  plastics |>
    filter(country == input_country) |>
    distinct(num_events, volunteers) |>
    mutate(
      total_vol = sum(volunteers, na.rm = TRUE),
      total_events = sum(num_events, na.rm = TRUE)
    ) |>
    distinct(total_vol, total_events) |>
    mutate(vol_per_event = total_vol / total_events) |>
    pull(vol_per_event)
}

volunteer_summary <- plastics |>
  distinct(country) |>
  mutate(
    avg_participation = map_dbl(country, cleanup_totals),
    rounded_participation = round(avg_participation)
  ) |>
  slice_max(avg_participation, n = 10)

volunteer_summary
```
